---
title: "Statistical report for ACASA-TAVI"
author: "C. Cunen"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: '2'
  #word_document:
  #  toc: true
  #  toc_depth: '2'
  #pdf_document:
  #  toc: true
  #  toc_depth: '2'
  #html_document:
  #  toc: true
  #  toc_depth: 2
  #  number_sections: true
subtitle: AntiCoagulation versus AcetylSalicylic Acid after Transcatheter Aortic Valve Implantation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr) #for tables
library(flextable) # for nice tables
library(broom) # for tidy function 

options(tidyverse.quiet = TRUE) # suppress package loading messages
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) #to set the working directory back to the folder containing the .Rproj file during knitting
```

```{r, message = TRUE}
rm(list = ls())

#datafile = '../data/raw/UiO_20250821_115105.xlsx'
#treatment_file = '../data/raw/2025-08-20 BASIC avblinding.xlsx'
#if (grepl('Fake', treatment_file, fixed = TRUE) == TRUE | grepl('fake', treatment_file, fixed = TRUE) == TRUE) {
#print('We are using a fake allocation list!')
#}

```

```{r, message = FALSE, warning=FALSE}
#source('../Rcode/raw_to_td.R')
#source('../Rcode/tables_descriptive.R') # a script with many tables
#source('../Rcode/raw-to-ad.R')
load("data/res/disp_tab.RData")
load("data/res/baseline_tab.RData")
```

# Introduction

The ACASA-TAVI study is ~~a prospective investigation of losartan as an adjunctive therapy for patients with glioblastoma~~. According to Protocol ~~Version 7.0 (17-Dec-2024), the trial collects comprehensive demographic, clinical, imaging, and quality-of-life information prior to treatment initiation~~. 

This report summarises ~~the currently available baseline data with emphasis on clinical status, imaging assessments, and prior treatments that may influence subsequent outcomes~~. 

 The safety analysis is based on 
the Safety Analysis Set including all subjects with any safety information after baseline (N=53). 
Patients randomised to hydroxychloroquine without receiving any amount of the treatment will be 
excluded from the safety population (nobody was excluded based on that condition). The primary 
efficacy analysis is based on the Full Analysis Set (FAS) including all randomised subjects who have 
had at least one baseline and one post-randomisation evaluation of efficacy (N=51). The secondary 
efficacy analysis is based on the Per Protocol Set including all participants of the FAS who took at 
least 7 of the 8 first hydroxychloroquine doses (N=49) 
Analysis according to final SAP version 2.1, dated 28.05.2020 
Data extraction Viedoc electronic data capture system, time stamped: 26.05.2020 12:06 UTC 

The aim of this document is to produce tables and figures to be included into the article as requested by email from Merete Hetland 10 December 2019. The analyses are prespecified in the Statistical Analysis Plan (SAP) dated 2nd of September 2019, and the locked database with an update received on the 10th of December dated 26th of October 2019.


## Randomisation Status

The present dataset is built on *sham randomisation* generated from the most recent export from Viedoc. No true treatment assignments have been unlocked. All findings in this document should therefore be interpreted as preliminary and descriptive only. Final efficacy comparisons will not be conducted until the database is locked and genuine randomisation codes are used.

## Data preparations

The data have been further prepared in accordance with the SAP, except for the imputation of missing baseline Body Mass Index (BMI) for two trial subjects. They have been imputed with the median BMI value among all other ITT subjects. This imputation is not described in the SAP.  

## Specifications to the analyses

For the continuous CRP and ESR parameters, I have fitted a generalized linear mixed gamma model due to skewness. This model gave the best fit according to Akaike's Information Criterion over a standard linear mixed model and a generalized linear mixed lognormal model. For the joint count variables I fitted a generalised linear mixed negative binomial model over a corresponding poisson model due to overdispersion. All these specification are according to the analysis plan. 

# Patient enrolment and disposition 

## Trial population 

The following table summarizes the number of patients that were screened (?), that were eligible and randomized, that received the treatment and that were lost to follow-up, as well as the number of patients included in the analyses. This table should be used to fill in the CONSORT diagram (??). 

```{r}
# consort_tab %>% 
#   head(10) %>%
#   kable(
#     caption = "Trial population",
#     digits = 2
#   )
tt <- flextable(consort_tab)
autofit(tt) %>%
  set_header_labels(values=list(row_label=" ")) %>%
  set_caption(caption = "Counts")
```

## Patient disposition (reasons for non-completion)

Overview of study completion and reasons for non-completion. The counts of "Study completed" which are neither "Yes" nor "No" are participants who have not had their final visit yet.

```{r}
body_tab <- body_tab %>% mutate(Total=DOAC+ASA)
tt <- flextable(body_tab)
tt %>%
  colformat_char(j = 1, na_str="NA") %>%
  flextable::compose(
    i = c(3:4,6:10),
    j = "row_label",
    value = as_paragraph("\t", row_label)) %>%
  flextable::compose(
    i = c(1:2,5),
    j = "row_label",
    value = as_paragraph(as_b(row_label))) %>%
  set_header_labels(values=list(row_label=" ")) %>%
  autofit() %>%
  set_caption(caption = "Overview of study completion and reasons for non-completion")

```

## Timing of end of study, withdrawals and follow ups


Time from randomisation to end of study and time from randomisation to withdrawal/lost to follow-up is summarized by median and min-max, overall and by treatment group. 

```{r}
tt <- flextable(time_tab)
tt %>%
  set_header_labels(values=list(dsc=" ")) %>%
  autofit() %>%
  set_caption(caption = "Length of study by treatment groups")
```


## Protocol deviations

Number of major and minor protocol deviations, summarized by treatment group with details of type of deviation provided. 

```{r}
tt <- flextable(minorPD)
tt %>%
  set_header_labels(values=list(ran_trt="Treatment",cat="Category of PD")) %>%
  autofit() %>%
  set_caption(caption = "Number of minor protocol deviations")
```

There are currently no PD that are categorised as "major".

```{r}
tt <- flextable(majorPD)
tt %>%
  set_header_labels(values=list(ran_trt="Treatment",cat="Category of PD")) %>%
  autofit() %>%
  set_caption(caption = "Number of major protocol deviations")
```


## Adherence to allocated treatment 


Number of patients participants switching between ASA and DOAC or stopping the treatment altogether (so far - final data could reveal more switches), by treatment group and in total.

```{r}
tt <- flextable(adh_tab)
tt %>%
  set_header_labels(values=list(ran_trt=" ",n_noChange="No change in study drug",
                                n_switch="Switching",n_stop="Stopping")) %>%
  autofit() %>%
  set_caption(caption = "Number of participants switching or stopping.")
```





## Analysis sets

The FAS includes all participants. The SAS is the same as the FAS.
The PP includes all participants except 
-- non-completers that did not die (voluntary discontinuation mostly)
-- participants with important protocol deviations
-- participants switching between ASA and DOAC or stopping the treatment altogether.


```{r}
tt <- flextable(set_tab)
tt %>%
  set_header_labels(values=list(ran_trt=" ",FAS="Full analysis set",PP="Per protocol set")) %>%
  autofit() %>%
  set_caption(caption = "Number of participants in the different analysis sets.")
```


# Demographics and baseline characteristics


## Table for abstract (May 2026)


```{r}
tt <- flextable(dm_table_abs)
tt %>% 
  set_header_labels(values=list(text=" ","1"=" ")) %>%
  autofit() %>%
  set_caption(caption = "Baseline characteristics for all 360 participants. Continuous variables are summarised by mean (SD), or by median (IQR) as indicated.",align_with_table = FALSE)

```


## Demographics and medical history

```{r}
indent <- c(3:4,6:7,9:10,12:13,15:16,18:19,21:22,24:27)
notindent <- which(!1:27 %in% indent)
tt <- flextable(dm_baseline)
tt %>%
  flextable::compose(
    i = indent,
    j = "Characteristic",
    value = as_paragraph("\t", Characteristic)) %>%
  flextable::compose(
    i = notindent,
    j = "Characteristic",
    value = as_paragraph(as_b(Characteristic)))  %>%
  autofit() %>%
  set_caption(caption = "Demographics and medical history.",
              align_with_table = FALSE)

```

## Vital signs and physical characteristics at baseline

```{r}
indent <- c(10:12,14:16,18:21)
notindent <- which(!1:21 %in% indent)
tt <- flextable(ph_tab)
tt %>%
  flextable::compose(
    i = indent,
    j = "Characteristic",
    value = as_paragraph("\t", Characteristic)) %>%
  flextable::compose(
    i = notindent,
    j = "Characteristic",
    value = as_paragraph(as_b(Characteristic)))  %>%
  autofit() %>%
  set_caption(caption = "Vital signs and physical characteristics at baseline.",
              align_with_table = FALSE)

```

## Clinical chemistry and hematology at baseline

```{r}
indent <- c(14,20,22)
notindent <- which(!1:23 %in% indent)
tt <- flextable(om_baseline)
tt %>% flextable::compose(
    i = indent,
    j = "Characteristic",
    value = as_paragraph("\t", Characteristic))  %>%
  autofit() %>%
  set_caption(caption = "Some results related to clinical chemistry and hematology at baseline.",align_with_table = FALSE)

```


## Echocardiography pre-TAVI and at baseline

```{r}
tt <- flextable(eco_screening)
tt  %>%
  autofit() %>%
  set_caption(caption = "Some echocardiography parameters pre-TAVI.",align_with_table = FALSE)

```

```{r}
tt <- flextable(eco_baseline)
tt  %>%
  autofit() %>%
  set_caption(caption = "Some echocardiography parameters at baseline (post-TAVI).",align_with_table = FALSE)

```

## Characteristics of the TAVI procedure

```{r}
indent <- c(2:3,5:6,8:9,11)
notindent <- which(!1:11 %in% indent)
tt <- flextable(tavi_baseline)
tt %>% flextable::compose(
    i = indent,
    j = "Characteristic",
    value = as_paragraph("\t", Characteristic)) %>%
  flextable::compose(
    i = notindent,
    j = "Characteristic",
    value = as_paragraph(as_b(Characteristic)))  %>%
  autofit() %>%
  set_caption(caption = "Characteristics of the TAVI procedure.",align_with_table = FALSE)

```

## Change in baseline characteristics

> Some baseline variables are measured again at 12 months. Do you want a table looking at the change in these values?
> Not here maybe


# Primary endpoints


## Primary efficacy endpoint

### Descriptive statistics

Also by site?

```{r}
load("data/res/prim_eff_tab.RData")
tt <- flextable(effsum2)
tt %>%
  set_header_labels(values=list(cthalt=" ")) %>%
  flextable::compose(
    j = "cthalt",
    value = as_paragraph(as_b(cthalt)))  %>%
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint). Counts and proportions.",align_with_table = FALSE)

```

### Main analysis: Unadjusted analysis with multiple imputation

Multiple imputation with imputation model containing sex, hemoglobin, potassium, 
bmi, sodium and LDL cholesterol. So the model does not use the treatment variable 
(to stay more or less in-line with the protocol).
But the analysis model only uses the treatment variable and no other covariates.

```{r}
halt_main$contrast <- "DOAC / ASA"
tt <- flextable(halt_main[,c("contrast","estimate","p.value",
                            "conf.low","conf.high")])
tt %>%
  set_header_labels(values=list(conf.low="95CI_low",conf.high="95CI_high")) %>% 
  colformat_double(j = 2:5, digits = 3) %>%
  autofit() %>%
  set_caption(caption = "Estimated average unadjusted risk ratio, along with corresponding p-value and confidence interval.",align_with_table = FALSE)

```


### Sensitivity analyses

Unadjusted analysis, complete-cases

Dropping participants with missing HALT value (ie. dropping `r length(summary(mod0)$na.action)` observations - remember that some participants with missing HALT are dead and so are counted as 'events' in this analysis).

```{r}
#tt <- flextable(comp0_cc[,c("contrast","estimate","std.error","p.value",
#                            "conf.low","conf.high")])
#tt %>%
#  set_header_labels(values=list(std.error="SE",conf.low="95CI_low",conf.high="95CI_high")) %>% 
#  colformat_double(j = 2:6, digits = 3) %>%
#  autofit() %>%
#  set_caption(caption = "Estimated average unadjusted risk difference, along with corresponding p-value and confidence interval.",align_with_table = FALSE)

```

Adjusted analysis with multilple imputation

Multiple imputation with imputation model containing sex, hemoglobin, potassium, 
bmi, sodium and LDL cholesterol. So the model does not use the treatment variable 
(to stay more or less in-line with the protocol).
The analysis model uses the covariates listed above and the treatment variable in addition.


```{r}
# tt <- flextable(comp1_mi[,c("contrast","estimate","std.error","p.value",
#                             "conf.low","conf.high")])
# tt %>%
#   set_header_labels(values=list(std.error="SE",conf.low="95CI_low",conf.high="95CI_high")) %>% 
#   colformat_double(j = 2:6, digits = 3) %>%
#   autofit() %>%
#   set_caption(caption = "Estimated average adjusted risk difference, along with corresponding p-value and confidence interval.",align_with_table = FALSE)

```



Adjusted analysis, complete-cases

Dropping participants with missing HALT value (ie. dropping `r length(summary(mod1)$na.action)` observations - remember that some participants with missing HALT are dead and so are counted as 'events' in this analysis).

```{r}
# mod_sum1 <- tidy(mod1)
# tt <- flextable(mod_sum1[,-4])
# tt %>%
#   set_header_labels(values=list(term=" ",std.error="SE")) %>% 
#   colformat_double(j = 2:4, digits = 2) %>%
#   autofit() %>%
#   set_caption(caption = "Fitted model.",align_with_table = FALSE)

```


```{r}
# tt <- flextable(comp1_cc[,c("contrast","estimate","std.error","p.value",
#                             "conf.low","conf.high")])
# tt %>%
#   set_header_labels(values=list(std.error="SE",conf.low="95CI_low",conf.high="95CI_high")) %>% 
#   colformat_double(j = 2:6, digits = 3) %>%
#   autofit() %>%
#   set_caption(caption = "Estimated average adjusted risk difference, along with corresponding p-value and confidence interval.",align_with_table = FALSE)

```




## Primary safety endpoint

### Descriptive statistics


```{r}
load("data/res/prim_saf_tab.RData")
load("data/res/prim_saf_fig.RData")
tt <- flextable(safsum)
tt %>%
  set_header_labels(values=list(event_type=" ")) %>%
  flextable::compose(
    j = "event_type",
    value = as_paragraph(as_b(event_type)))  %>%
  autofit() %>%
  set_caption(caption = "Descriptive statistics for primary safety endpoint. Counts and proportions. The counts are the number of participants experiencing an event of a particular type as their *first* safety event. Some patients may experience more than one safety event.",align_with_table = FALSE)

```

Out of the `r saf_events$total` safety events, `r saf_events$saf_switch` happen in patients that also switch between the treatments at some point. Among these  `r saf_events$saf_switch` patients:  `r saf_events$before` experience the safety event before switching, while  `r saf_events$after` experience the safety event after switching.


```{r, echo = FALSE, fig.width = 12, fig.height=6, fig.cap="All patients"}

all

```


```{r, echo = FALSE, fig.width = 12, fig.height=8, fig.cap="Patients with safety events and/or switching and discontinuation"}
load("data/res/prim_saf_fig.RData")
selected

```

In this figure there are 46 switches and 7 discontinuations, and these number may seem to bit slightly at odds with the numbers presented in Table 2.6. But that's actually just because what is called "switches" here in fact includes both "switches and stops" in Table 2.6.

### Unadjusted analysis with multiple imputation, treatment policy strategy

Multiple imputation with imputation model containing no covariates
(so using only the overall event rate, to stay more or less in-line with the protocol).
The analysis model uses the treatment variable and no other covariates.

```{r}
tt <- flextable(scomp0_mi[,c("contrast","estimate","std.error","p.value.nonsup",
                            "conf.low","conf.high")])
tt %>%
  set_header_labels(values=list(std.error="SE",conf.low="95CI_low",conf.high="95CI_high")) %>% 
  colformat_double(j = 2:6, digits = 3) %>%
  autofit() %>%
  set_caption(caption = "Estimated average unadjusted risk difference, along with corresponding p-value and confidence interval.",align_with_table = FALSE)

```


### Unadjusted analysis, complete-cases, treatment policy strategy

Discuss margin (in protocol it says 11.9%, but that seems a bit wide maybe)


```{r}
tt <- flextable(scomp0_cc[,c("contrast","estimate","std.error","p.value.nonsup",
                            "conf.low","conf.high")])
tt %>%
  set_header_labels(values=list(std.error="SE",conf.low="95CI_low",conf.high="95CI_high")) %>% 
  colformat_double(j = 2:6, digits = 3) %>%
  autofit() %>%
  set_caption(caption = "Estimated average unadjusted risk difference, along with corresponding p-value (for non-inferiority test) and confidence interval.",align_with_table = FALSE)

```

# Key secondary endpoints

## Clinical efficacy

## Thromboembolic events

## Bleeding events

## All-cause mortality



# Secondary safety endpoints

# Sub-group analyses

## HALT

### Age


```{r}
load("data/res/subgroups_prim_eff_tab.RData")
tt <- flextable(ageT[3:6,])
tt %>%
  add_header_row(values=c(" ","<75",">= 75"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) in age subgroups. Counts and proportions.",align_with_table = FALSE)

```

### Sex


```{r}
tt <- flextable(sexT[3:6,])
tt %>%
  add_header_row(values=c(" ","Male","Female"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by sex. Counts and proportions.",align_with_table = FALSE)

```


### Diabetes


```{r}
tt <- flextable(diabT[3:6,])
tt %>%
  add_header_row(values=c(" ","No diabetes","Diabetes"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by diabetes. Counts and proportions.",align_with_table = FALSE)

```


### Hypertension


```{r}
tt <- flextable(hypT[3:6,])
tt %>%
  add_header_row(values=c(" ","No hypertension","Hypertension"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by hypertension. Counts and proportions.",align_with_table = FALSE)

```


### Impaired renal function

```{r}
tt <- flextable(renT[3:6,])
tt %>%
  add_header_row(values=c(" ","Impaired","Preserved"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by renal function. Counts and proportions.",align_with_table = FALSE)

```

> Here there are so few in the impaired group that I don't think inference is necessary.


### TAVI valve type

```{r}
tt <- flextable(valT[3:6,])
tt %>%
  add_header_row(values=c(" ","Balloon-expanded","Self-expandable"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by valve type. Counts and proportions.",align_with_table = FALSE)

```


### TAVI post-dilation

```{r}
tt <- flextable(posT[3:6,])
tt %>%
  add_header_row(values=c(" ","No","Yes"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by post-dilation. Counts and proportions.",align_with_table = FALSE)

```

### TAVI supra-valcular prosthesis position

```{r}
tt <- flextable(supT[3:6,])
tt %>%
  add_header_row(values=c(" ","No","Yes"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by supra-valcular prosthesis position. Counts and proportions.",align_with_table = FALSE)

```

### Ascending aorta diameter

```{r}
tt <- flextable(ascT[3:6,])
tt %>%
  add_header_row(values=c(" ","< 30",">= 30"),colwidths = c(1,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by ascending aorta diameter (mm). Counts and proportions.",align_with_table = FALSE)

```

### Frailty index

```{r}
tt <- flextable(fraT[3:6,])
tt %>%
  add_header_row(values=c(" ","robust","pre-frail","missing"),colwidths = c(1,2,2,2),top=T)  %>%
  set_header_labels(values=list(at=" ",V1="DOAC",V2="ASA",V3="DOAC",V4="ASA",V5="DOAC",V6="ASA"))  %>% 
  autofit() %>%
  set_caption(caption = "Descriptive statistics for HALT (primary efficacy endpoint) by frailty category. Counts and proportions.",align_with_table = FALSE)

```


# Additional safety analysis

> Which numbers and tables do you want to see here? The two first tables are kind of redundant, but the question is whether one should require that the AE is possibly related to study medication or not.

## Overview of adverse events, serious adverse events and suspected unexpected serious adverse reaction


```{r}
load("data/res/ae_tab.RData")
tt <- flextable(ae_table)
tt %>%
  set_header_labels(values=list(text=" "))  %>%
  autofit() %>%
  set_caption(caption = "Summary of adverse events.",align_with_table = FALSE)

```

There are no SUSARs in the dataset. No SAE which is unexpected and possibly, probably or definitely related to study medication.

## Suspected adverse reactions: AEs and SAEs with a possible, probable or definite relation to study treatments


```{r}
tt <- flextable(pae_table)
tt %>%
  set_header_labels(values=list(text=" "))  %>%
  autofit() %>%
  set_caption(caption = "Summary of adverse events with a possible, probable or definite relation to study treatments.",align_with_table = FALSE)

```

There are no SUSARs in the dataset.

## Adverse Events by System Organ Class and Preferred term 

```{r}
tt <- flextable(ae_sum)
tt %>%
  add_header_row(values=c(" ","AE","SAE"),colwidths = c(2,2,2),top=T)  %>%
  set_header_labels(values=list(soc="System organ class",pt="Preferred term","0_DOAC"="DOAC",
                                "0_ASA"="ASA","1_DOAC"="DOAC","1_ASA"="ASA"))  %>%
  autofit() %>% theme_box %>%
  set_caption(caption = "Number of AEs and SAEs by system organ class and preferred term.",align_with_table = FALSE)

```